---
title: "Markov Chains in R"
author: "Gilbert"
date: "October 2016"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Markov Chains in R

The package markovchain in R can be used to work with Markov Chains. We can load it and define a markov chain as follows:

```{r cars}
library(markovchain)
tmA <- matrix(c(0,0.5,0.5,.5,0,.5,.5,.5,0),nrow = 3,
byrow = TRUE) #define the transition matrix

dtmcA <- new("markovchain",transitionMatrix=tmA,
states=c("a","b","c"),
name="MarkovChain A") #create the DTMC

dtmcA
```

## Including Plots

You can also plot a display with the igraph, for example:

```{r pressure, echo=FALSE}
plot(dtmcA)
```

## Transistion probabilities

The easiest way to retrieve a transition probability and conditional distribution is:
``` {r tp}
dtmcA[2,3]
conditionalDistribution(dtmcA,"b")
```

We can also simulate states distribution after n-steps, and the steady state distributions:

```{r f}
initialState<-c(0,1,0)
steps<-4
finalState<-initialState*dtmcA^steps #using power operator
finalState

steadyStates(dtmcA)
```

## Another Example:

```{r g}
E <- matrix(0, nrow = 4, ncol = 4)
E[1, 2] <- 1;E[2, 1] <- 1/3; E[2, 3] <- 2/3
E[3,2] <- 1/4; E[3, 4] <- 3/4; E[4, 3] <- 1
MC <- new("markovchain", states = c("a", "b", "c","d"), transitionMatrix = E,name = "MC")
summary(MC)
```

This is a good environment to work with DTMCs



