---
title: "Analysis of mtcars dataset"
output:
  pdf_document: default
  html_document: default
---
### Executive Summary ###

The dataset mtcars contains data that was extracted from the 1974 Motor Trend US magazine, and comprises fuel consumption and 10 aspects of automobile design and performance for 32 automobiles (1973-74 models). Here we analyze the dataset to see if any significant relationships exist between Miles per Gallon(mpg) and any of the other recorded variables.  We are especially interested in whether or not there is a difference in mpg for automatic versus manual transmissions, and if so which is better.

### Data Analysis ###
We begin by looking at the structure of the dataset.  Plots and variable information are contained in the appendix.  We have 11 variables, including mpg, which will be our response variable in what follows, and am, which is coded as 0 for automatic vehicles and 1 for manual.  First we will fit a linear model with all variables, and then use stepwise backward regression to find the "best" model:

```{r}
full.model <- lm(mpg ~ .,mtcars)
summary(full.model)
```
INterestingly, although the F statistics is significant for the model, not variables is significant at the 5% level, although wt is close.

Now the reduced model is fit, detailed output of which can be found in the appendix:
```{r, results='hide'}
reduced.model <- step(full.model, direction="backward")
```
and summary:
```{r}
summary(reduced.model)
```

This model keeps quarter mile time (qsec), weight (wt), and am, all significant at the 5% level. 
Here are the confidence intervals;

```{r}
confint(reduced.model)
confint(reduced.model, level=0.99)
```

As expected, all three variables have C.I.'s that do not contain 0 at the 95% level.  However the 99% CI for am does contain 0, implying it is not significant at the 99% level.

Next, we examine the residual plot:
```{r}
plot(reduced.model, which=1)
```
Nothing suspicious show up here.

### Conclusion ###
We conclude that weight, quarter mile time, and automatic transmission vs manual are all significant at the 5% level, and only the first 2 at the 99% level.  The estimate of 2.9 for the am variable implies that manual transmission will give a boost of 2.9 mpg over automatic.  However, with only 32 automobiles in the dataset and it not being significant at the 99% level, we advise noting the uncertainty of the results. 

### Appendix ###
Exploratory data analysis of mtcars:
```{r}
str(mtcars)
require(graphics)
pairs(mtcars, main = "mtcars data")
coplot(mpg ~ disp | as.factor(cyl), data = mtcars,
       panel = panel.smooth, rows = 1)
boxplot(mpg ~ as.factor(am), data=mtcars, main = "Boxplot of MPG by Automatic(0) and Manual(1)")
```

Data from the backward regression:
```{r}
reduced.model <- step(full.model, direction="backward")
```
